---
title: "What is the oldest property in Montreal?"
subtitle: "From the series: real estate in Montreal"
author: Andryas Wavrzenczak
date: '2022-05-29'
slug: real-estates-year-built
categories:
  - real-estates
tags:
  - real-estates
  - montreal
  - year-built
comments: true
showTags: true
showPagination: true
showSocial: true
showDate: true
draft: false
thumbnailImage: https://mspublic.centris.ca/media.ashx?id=ADDD250D253D373D1B5EA62DE0&t=pi&sm=m&w=1024&h=1024
thumbnailImagePosition: left
summary: "Something exciting about the real estate data in Canada is that most of the  websites give the year built of each property, and through this, we will see  where the oldest part of Montreal is. Is the old port really old?"
editor_options: 
  chunk_output_type: console
---

```{r setup, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  fig.height = 9,
  fig.width = 11
)
```

```{r libs}
library(sf)
library(rgdal)
library(tidyverse)
library(echarts4r)
library(geojsonsf)
library(kableExtra)
library(wesanderson)

library(ggsflabel)

library(aw)

prop <- readRDS("data/properties.rds")
prop$tipology2 <- ifelse(prop$tipology == "house", "house", "condo")
prop <- prop |>
  select(url, year_built, lat, lng, tipology2, neighbourhood)
montreal <- readOGR("data/limadmin-shp/LIMADMIN.shp", verbose = FALSE) |>
  st_as_sf(crs = 4326)
montreal_json <- geojsonsf::sf_geojson(montreal |> rename(name = NOM))
```


Before start is essential to say that what we are exploring here is data crawled
from the Centris, so it is a property sample from Montreal; my inferences here 
will be estimation, approximation for the true values; it is good to give an 
idea about the places in Montreal.

Let's start seeing a summary of the whole data.

```{r}
simple_summarize(data = prop, var = "year_built") |>
  select(-var) |>
  kbl(digits = 4, caption = "Summary year built") |>
  kable_styling(c('striped', 'hover')) |>
  scroll_box(width = "100%")
```

The table above shows that 50% of the properties we collected were built before 
1992 (med=1992). The oldest house we found is dated 1690! I don't know if it is 
correct, but if it is, congratulation to Montreal; you've been pretty strict in 
dating properties.

Now let's open it by neighbourhood.

```{r}
simple_summarize(data = prop, var = "year_built", neighbourhood) |>
  select(-var) |>
  arrange(min) |>
  kbl(digits = 4) |>
  kable_styling(c('striped', 'hover')) |>
  scroll_box(width = "100%")
```

As we can see from the table above, the oldest property we have in our database
comes from Ville-Marie (1690); besides, it is the neighbourhood with the most 
significant number of observations, 2358.

A good visualition for this kind situation is the boxplot.

```{r}
prop |>
  ggplot(aes(x = neighbourhood, y = year_built)) +
  geom_boxplot() +
  coord_flip() +
  theme_minimal()
```

We can see the year-built distribution by each neighbourhood.

A visualization of all properties that we collected from centris. We notice 
that it is predominantly from 1900 later, but let's break the year into classes.

```{r}
prop |>
  drop_na(year_built) |> 
  ggplot() +
  geom_sf(data = montreal) +
  geom_point(aes(x = lng, y = lat, color = year_built)) +
  scale_colour_gradientn(colours = wes_palette("Zissou1", 5, type = "continuous")) +
  theme_void() +
  theme(legend.title = element_blank())
```

In the plot below, we can see better the patterns; first, the oldest properties 
in Montreal are in the old port, so the old port lives up to the name. Second, 
we have many properties between [1800-1900], and I can't stop wondering how 
they could have kept this information all these years. And third, the most 
common building age in our database from centris is between [1900-2000], 
followed by [2000-2022].  

```{r}
prop |>
  mutate(year_built_cut = cut(year_built, c(1600, 1800, 1900, 2000, 2022), dig.lab = 10)) |>
  drop_na(year_built) |>
  ggplot() +
  geom_sf(data = montreal) +
  geom_point(aes(x = lng, y = lat, color = year_built_cut)) +
  scale_color_viridis_d(option = "inferno") +
  theme_void() +
  theme(legend.title = element_blank()) +
  facet_wrap(~year_built_cut)
```

To end this post, a shortlist of the 10 oldest properties in Montreal (remember 
we crawled this data from centris, so it is a sample).

```{r}
prop |>
  arrange(year_built) |>
  slice(1:10) |>
  select(-lat, -lng,-tipology2) |> 
  kbl(digits = 4) |>
  kable_styling(c('striped', 'hover'))
```

Cheers.